Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-08-12 Latex_files/Thesis/chapters/chapter 1/main.tex
updates to ch3

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase separation 
used by humanity\cite{Schoen1956}, put simply, it is the 
formation and growth of a new structured phase within 
a disordered bulk phase. This has applications in a 
number of industries such as pharmaceuticals, food 
production, and electronics \cite{Myerson2002}; where 
the extraction of dilute materials can help improve 
product quality while keeping production costs low. 
The industrialisation of crystallisation has allowed 
engineers to reliably and efficiently induce the crystal 
formation within a bulk phase. However, while on a 
large scale crystallisation is seemingly an understood 
physical process, just a small amount of investigation 
into the literature reveals that at a micro scale there 
is yet to be unifying theory that can accurately explain 
the process of crystallisation \cite{Fu2021}. Of which 
there two primary areas of focus, nucleation and crystal 
growth. The latter focusing on how an already stable 
crystal grows and how it takes on its final shape; 
whereas the former is more concerned with which factors 
contribute to its formation in order to control crystal growth.

The focus of this research is partly based around investigating
the impact of a focused electromagnetic field can influence the 
nucleation process by inducing shear flow within the bulk phase. 
This in hope that the localised EM field can be used to study 
different phenomena in nucleation in order to develop a better
understanding of the underlying mechanisms.

@@

Where $C_{sol}$ is just the solute concentration, $C_{eq}(T)$
is the equilibrium concentration at temperature $T$.
While the solution remains supersaturated there is a 
chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, 
the first formation of the crystal is referred to as the 
nucleus and understanding its formation of has been the 
focus of researchers for decades now. Typically, for an 
industrial crystallisation process the working principle 
is based on controlling and manipulating the supersaturation
of the system.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can
be broadly categorised into either primary of secondary 
nucleation. The former being described as the formation of an
initial nucleus within the bulk phase; absent of any external
stimuli primary nucleation is considered stochastic and is typically
undesirable for industrial applications where short and consistent 
residence times are desirable. At a small scale one can estimate the 
nucleation rate by making repeated measurements of sample solutions 
and seeing how many have nucleated after a given time, giving us a 
Poisson distribution.

@@

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$
is the volume of the individual samples, and $M^*(t)$ \& $M$ are the 
number of nucleated samples and the total number of samples used respectively.
While this is useful for studying the effects of different parameters
at a small scale, for industrial applications there are so many
additional factors at play that can throw off this estimation.

@@

\section{Nucleation Theories}
\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the 
original theory was first formed from the works of Volmer 
and Weber, and Frenkel \cite{Frenkel1939, Volmer1926}. 
While initially it was more focused into describing droplet 
formation in condensing vapours it was extrapolated to describe
crystallisation. The central premise of classical theory 
is that nucleation occurs stochastically throughout the bulk
phase due to collisions between individual solute molecules,
ions, or atoms; at the same time the bulk phase is resistant
to the formation of a new phase. The competition between
these random collisions and the bulk solution can be used to 
predict the probability of a newly formed nucleus.

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. 
The newly formed phase has a lower chemical potential than the 
surrounding solution, reducing the free energy of the system; 
at the same time, the formation of a new interface is resited 
by the bulk phase due to surface tension. The net free energy 
of the system for a nucleus of radius $r$ is given as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf} 
\end{align}

Where $v$ is the approximate volume of an individual molecule, $k_B$
is the Boltzmann constant, and $\sigma_{inf}$ is the interfacial 
tension of the bulk solution. Plotting the free energy of the system
against nucleus size reveals a critical size above which the gain 
in free energy exceeds the interfacial tension. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	     \label{fig:free_energy}
\end{figure}

@@

\subsection{Optical Rotation}
For any electromagnetic field it is possible to transfer both linear and 
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

@@

\subsection{Shear induced Nucleation}
It has long been known that fluid shear rate plays a role in influencing
nucleation; however, the exact relationship between shear rate and nucleation
rate has only been recently understood for specific solutions. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

>>>>

\chapter{Introduction}
Crystallisation is one of the oldest forms of phase separation used by humanity\cite{Schoen1956}, put simply, it is the formation and growth 
of a new structured phase within a disordered bulk phase. This has 
applications in a number of industries such as pharmaceuticals, food 
production, and electronics \cite{Myerson2002}. Where the extraction 
of dilute materials can help improve product quality while keeping 
production costs low. The industrialisation of crystallisation has 
allowed engineers to reliably and efficiently induce the crystal 
formation within a bulk phase. However, while on a large scale 
crystallisation is seemingly an understood physical process, just a 
small amount of investigation into the literature reveals that at a 
micro scale there is yet to be unifying theory that can accurately 
explain the process of crystallisation \cite{Fu2021}. Of which there 
two primary areas of focus, nucleation and crystal growth. The latter 
focusing on how an already stable crystal grows and how it takes on its 
final shape. Whereas the former is more concerned with which factors 
contribute to its formation in order to control crystal growth.

The focus of this research is partly based around investigating the impact 
of a focused electromagnetic (EM) field can influence the nucleation process 
by inducing shear flow within the bulk phase. This in hope that the 
localised EM field can be used to study different phenomena in nucleation 
in order to develop a better understanding of the underlying mechanisms.

@@

Where $C_{sol}$ is just the solute concentration, and $C_{eq}(T)$is the 
equilibrium concentration at temperature $T$. While the solution remains 
supersaturated there is a chemical potential driving force for the solute 
to coalesce and separate from the solution as an ordered solid, the first 
formation of the crystal is referred to as the nucleus and understanding its 
formation of has been the focus of researchers for decades now. Typically, 
for an industrial crystallisation process the working principle is based on 
controlling and manipulating the supersaturation of the system. Of course 
nucleation events in a variety of different situations which each require 
their own classification.

\subsection{Primary \& Secondary nucleation}
From an industrial perspective, the nucleation process can be broadly 
categorised into either primary of secondary nucleation. The former 
describes the formation of an initial nucleus within the bulk phase, 
absent of any external stimuli. Primary nucleation is therefore considered 
stochastic as there is no reliable means of predicting where a nucleus
may form, or how long it will take. The only reliable indicator being 
that higher supersaturations will result in faster nucleation rates. At 
a small scale one can estimate the nucleation rate by making repeated 
measurements of sample solutions and seeing how many have nucleated 
after a given time, giving us a Poisson probability distribution.

@@

Where $J$ is the nucleation rate, $t_g$ is the 'growth time', $V$ is 
the volume of the individual samples, and $M^*(t)$ \& $M$ are the number 
of nucleated samples and the total number of samples used respectively.
While this is useful for studying the effects of different parameters at 
a small scale, for industrial applications there are to many external 
factors for primary nucleation to be the only

@@

\section{Nucleation Theories}
\subsection{Classical Nucleation Theory (CNT)}
Sometimes referred to as 'Gibs Nucleation Theory' the original theory 
was first formed from the works of Volmer and Weber, and Frenkel 
\cite{Frenkel1939, Volmer1926}. While initially it was more focused 
into describing droplet formation in condensing vapours it was 
extrapolated to describe crystallisation. The central premise of 
classical theory is that nucleation occurs stochastically due to 
collisions between individual solute molecules, ions, or atoms. At the 
same time the bulk phase is resistant to the formation of a new phase. 
The competition between these random collisions and the bulk solution 
can be used to predict the probability of a newly formed nucleus.

Consider a supersaturated solution, after some time enough individual
sub units collide, forming a nucleus of volume $4\pi r^3/3$. The newly 
formed phase has a lower chemical potential than the surrounding solution, 
reducing the free energy of the system. Simultaneously, the formation of 
a new interface is resited by the bulk phase due to surface tension. 
The net free energy of the system for a nucleus of radius $r$ is given 
as \cite{Karthika2016}:
\begin{align}
	\Delta G = \frac{-4\pi r^3}{3v}k_BT\ln(S) + 4\pi r^{2}\sigma_{inf}
	\label{eq:CNT} 
\end{align}

Where $v$ is the approximate volume of an individual molecule, $k_B$ is 
the Boltzmann constant, and $\sigma_{inf}$ is the interfacial tension of 
the bulk solution. This assumes that the nucleus will have a spherical 
morphology so that the surface tension $\sigma_{inf}$ is a scalar value. 
Looking at \eqref{eq:CNT} suggests that there must be some critical size 
$r$ where the free energy gain from the nucleus exceeds the surface tension 
of the surrounding fluid. Plotting the free energy of the system against 
nucleus size reveals a critical size above which the gain in free energy 
exceeds the interfacial tension. Furthermore it shows how increasing the 
supersaturation of the system reduces said barrier. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{Free_Energy_Diagram.png}
	\caption{Free energy diagram of a newly formed nucleus according 
		     to the Classical Nucleation Theory. The total free energy (blue)
		     is due to the competition between the volume free energy gain
		     (green) and the interfacial free energy cost (orange). Dotted
		     lines are for a higher supersaturation than the solid lines,
		     the interfacial energy cost is independent of supersaturation.}
	     \label{fig:free_energy}
\end{figure}

@@

\subsection{Optical Torque}
For any electromagnetic field it is possible to transfer both linear and 
angular momentum \cite{Beth1936MechanicalDA}; more accurately the field 
is said to have both orbital and spin momentum. Though there is some 
debate on how to decompose the total momentum into these two components 
\cite{Bruce2020, Svak2018}, for this thesis we do not need to calculate 
the exact quantities and will instead look at the broader effects of both 
components. Orbital angular momentum arises from the shape of the wavefront 
of the particular field in question; for simple Gaussian beams the wavefronts 
are uniform and equally spaced resulting in the typical radiation pressure 
that Ashkin and co demonstrated \cite{Ashkin1980}. However, higher order 
modes of a Gaussian beam (for example: Laguerre-Gaussian modes) have 
non-uniform wave fronts meaning the orbital momentum has both angular and
linear components; depending on the relative size of the target particle 
one can induce rotation, or orbiting \cite{Bruce2020, Courtial2000}.

@@

\section{Shear induced Nucleation}
It has long been known that fluid shear rate plays a role in influencing
nucleation; however, the exact relationship between shear rate and nucleation
rate has only been recently understood for specific solutions. Theoretical research 
into shear induced nucleation suggests that there should be a slight increase 
in the nucleation rate at low shear rates, reaching a maximum increase in 
nucleation rate, and then at higher shear rates the nucleation rate begins 
to drop off. This has been shown theoretically for both simple colloidal
\cite{Mura2016,Debuysschere2023,Richard2015} and ice crystal formation 
\cite{Goswami2020}; however, no experimental work into these systems has 
been conducted to prove this is the case.  There is some experimental 
evidence for this phenomena in simple salt and protein solutions - though 
the authors emphasise that mechanical agitation cannot be ruled out - 
there has not been a exhaustive study into the shearing effects apart from 
in glycine solutions. In \cite{Debuysschere2023} it was found that a shear 
rate of around $3000\ s^{-1}$ was the maximum shear rate that would yield 
the highest nucleation rate. Using the theoretical model established in 
\cite{Mura2016,2001} which modifies the CNT to account for the effects of 
a nucleus undergoing shearing, accounting for the fact that a nucleus' 
growth is undergoing competition between flow-mediated molecular transport 
and the strain applied by the flow field which inhibits the growth of the 
nucleus. There central conclusion (from both the theoretical and experimental 
results) is that there is an optimal shear rate in which the nucleation 
rate is maximised. However, a question that arises from this result, if 
there is a optimal shear rate in which molecular transport is maximised 
and strain is minimised, then surely there should also be a shear rate 
in which the molecular transport and strain are equal - allowing one to 
suspend a nucleus at a constant radius. In this scenario, the molecular 
transport would prevent the nucleus from dissolving, but the strain would 
prevent the nucleus from growing. This however would require one to be 
able to apply a continuous shear rate to a targeted nucleus with high 
precision, there is also no model for an individual nucleus undergoing growth.

### 2024-08-12 Latex_files/Thesis/chapters/chapter 3/main.tex
updates to ch3

Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
other polymorphs of calcite and aragonite, when synthesised vaterite will 
typically form small spherical particles making them ideal for optical 
trapping and rotation. Synthesis of vaterite micro spheres requires fine 
control of the nucleation process in order to maintain polymorphic stability, 
though for the purposes of optical rotation the exact polymorph is not as 
important as its morphology as all 3 polymorphs are inherently birefringent.

@@

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical 
assumption being that the particle perfectly rotates in the $x-y$ plane; 
however it has long been suspected that birefringent microspheres 
experience torques outside of the $x-y$ plane \cite{Volpe2023} making 
it very difficult to characterise the behaviour of rotating birefringent 
microspheres without a proper understanding of the full optical torque 
being applied to it.

@@

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we computed the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field):
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around micro-rotors}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples were 
synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
\ref{sec:vaterite}), and then suspended in distilled water. A sample
of $200\ \mu L$ was pipptted and a single microsphere was captured via
a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck together, 
fortunately individual sphere's were still present. Multiple microspheres 
were trapped and their rotation rate was determined by looking at the 
peak frequency component of the collected power spectrum. The shear flow
was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
were operating in the bulk fluid and away from any boundaries.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.82\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
	\end{subfigure}
	\begin{subfigure}{0.82\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
	\end{subfigure}
	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
		vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
		is indicated by the dotted line. Vaterite radii and rotation frequencies 
		are shown, the laser power was kept constant at 450 mW. Reported rotation
		rates, and their corresponding fluid flow and shear rates, for vaterite 
		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
	\label{fig:vaterite_shear}
\end{figure}

From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
particle size and rotation rate, this is contrary to much of the theoretical
predictions that predict an exponential decay with particle size. This can
be in part due to the fact that synthesising perfectly spherical spheres that
have uniform birefringence across the whole population is difficult. Despite 
our best efforts at controlling the growth rate the smallest particle ever 
synthesised was around $3\ \mu m$ in diameter. The vaterite spheres would 
often stick together while suspended in water after a short period of time. 
The fastest reported rotation rate found during this project was by 
\cite{Arita2016} that achieved a rotation rate of $5\ kHz$, this is plotted 
on Fig~\ref{fig:vaterite_shear} as the dotted line. Even at that extreme a 
rotation rate the region in which nucleation is at its optimal likelihood is 
only $20\ nm$ wide. This suggests that a more complicated fluid flow 
arrangement is required in order to achieve the necessary shear rate and over 
a larger volume of fluid.

\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating birefringent 
		particles in a supersaturated solution. Left: side view of the 
		trapping set up showing the location of the trap focus at the 
		edge of the droplet of a supersaturated solution. Right: top down 
		view of the glycine droplet with a trapped birefringent particle 
		shown close to edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field around 
		itself (see Eq.~\ref{eq:birefringent_speed}).}
\end{figure}

After a period of ten minutes if no nucleation event was 
observed the particle was released and another particle was 
trapped. Due to the increased viscosity of the supersaturated 
solution and the proximity to the droplet edge, as previous works
have only seen nucleation near the solution-air interface \cite{Gowayed2021, 
Liao2022, Walton2018}, the observed rotation rate was rather low or 
non-existent, the results are tabulated below:
\begin{table}[h!]
	\centering
	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}

>>>>

Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
other polymorphs of calcite and aragonite, when synthesised vaterite will 
typically form small spherical particles making them ideal for optical 
trapping and rotation. Synthesis of vaterite micro spheres requires fine 
control of the crystal growth process in order to maintain polymorphic stability. 
Though for the purposes of optical rotation the exact polymorph is not as 
important as its morphology as all 3 polymorphs are inherently birefringent.

@@

As shown by Fig.~\ref{fig:vaterite}(b) the power spectra produced still 
demonstrates a Lorentzian curve but modified with these periodic peaks, 
while the Lorentzian can be loosely fitted to the end tail there exists 
no current model for describing the power spectra. The closest approximation
to this was conducted by \cite{Yogesha2012} where they describe the 
rotational motion of ellipsoidal polystyrene particles. The critical 
assumption being that the particle perfectly rotates in the $x-y$ plane. 
It has long been suspected that birefringent microspheres experience torques 
outside of the $x-y$ plane \cite{Volpe2023} making it very difficult to 
characterise the behaviour of rotating birefringent microspheres without a 
proper understanding of the full optical torque being applied to it.

@@

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field):
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around micro-rotors in bulk fluid}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples were 
synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
\ref{sec:vaterite}), and then suspended in distilled water. A sample
of $200\ \mu L$ was pipptted and a single microsphere was captured via
a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck together, 
fortunately individual sphere's were still present. Multiple microspheres 
were trapped and their rotation rate was determined by looking at the 
peak frequency component of the collected power spectrum. The shear flow
was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
were operating in the bulk fluid and away from any boundaries.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
	\end{subfigure}
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
	\end{subfigure}
	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
		vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
		is indicated by the dotted line. Vaterite radii and rotation frequencies 
		are shown, the laser power was kept constant at 450 mW. Reported rotation
		rates, and their corresponding fluid flow and shear rates, for vaterite 
		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
	\label{fig:vaterite_shear}
\end{figure}

From Fig.\ref{fig:vaterite_shear} there is not a strong relationship between
particle size and rotation rate, this is contrary to much of the theoretical
predictions that predict an exponential decay with particle size. This can
be in part due to the fact that synthesising perfectly spherical spheres that
have uniform birefringence across the whole population is difficult. Despite 
our best efforts at controlling the growth rate the smallest particle ever 
synthesised was around $3\ \mu m$ in diameter. The vaterite spheres would 
often stick together while suspended in water after a short period of time. 
The fastest reported rotation rate found during this project was by 
\cite{Arita2016} that achieved a rotation rate of $5\ kHz$, this is plotted 
on Fig~\ref{fig:vaterite_shear} as the dotted line. Even at that extreme a 
rotation rate the region in which nucleation is at its optimal likelihood is 
only $20\ nm$ wide. If instead the micro-rotor was within the vicinity of a
solid boundary, the shear rate would be enhanced due to the no-slip boundary
condition.

\section{Micro-rotors in Supersaturated solution}
If rotation rates in bulk solution are insufficient then a micro-rotor rotating
close to an artificial barrier may be able to improve the shear rate of the 
surrounding fluid. Of course placing a solid barrier in a supersaturated fluid
may well encourage nucleation somewhere on the surface outside of our control.
Instead we chose to use the droplet edge of the supersaturated solution, while
not a hard barrier per say, the molecular mobility close to the droplet edge is
reduced due to surface tension. Furthermore, it has been shown through multiple 
results that nucleation is enhanced at the air-solution interface \cite{Liao2022,
Yuyama2010, Sugiyama2009}. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{vaterite_diagram.pdf}
	\caption{Diagram of optical trapping set up for rotating birefringent 
		particles in a supersaturated solution. Left: side view of the 
		trapping set up showing the location of the trap focus at the 
		edge of the droplet of a supersaturated solution. Right: top down 
		view of the glycine droplet with a trapped birefringent particle 
		shown close to edge of the trap. As the particle rotates the drag 
		force from the surrounding fluid generates a flow field around 
		itself (see Eq.~\ref{eq:birefringent_speed}).}
\end{figure}

Supersaturated solutions of glycine and water were prepared and stored in
an incubator prior to use. When ready to be studied $15\ \mu L$ was pipetted 
into the solution and $20\ \mu L$ was pipetted onto the cover slip. A vaterite
sphere was located, trapped, and moved as close to the droplet edge as possible.
After measuring the microsphere's rotational frequency the sphere was left to
rotate for a period of ten minutes after which, if no nucleation event was 
observed the particle was released. The overall results are catalogued in 
table~\ref{tab:Nucleation}
\begin{table}[h!]
	\centering
	\caption{Results from rotating vaterite within supersaturated solution of $H_2O$ and Glycine. Solubility concentration for Glycine at $16^\circ$ was $C^*=0.2016g/g$}
	\label{tab:Nucleation}
	\begin{tabular}[width=\textwidth]{|c|c|c|c|}
		\hline
		Super Saturation & Particle radius [$\mu$] & $\omega$ [Hz] & Nucleation [$\checkmark/\times$]\\
		\hline
		\multirow{3}*{1.01} & 2.34 & 20.7 & $\times$ \\
		\cline{2-4} & 5.67 & 23.3 & $\times$ \\
		\cline{2-4} & 3.26 & 15.4 & $\times$ \\
		\hline
		\multirow{3}*{1.14} & 1.89 & 1.23 & $\times$ \\
		\cline{2-4} & 3.75 & 3.54 & $\times$ \\
		\cline{2-4} & 4.35 & 4.86 & $\times$ \\
		\hline
		\multirow{3}*{1.4} & 3.47 & 0.00 & $\times$ \\
		\cline{2-4} & 1.59 & 0.00 & $\times$ \\
		\cline{2-4} & 6.24 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.45} & 6.32 & 0.00 & $\times$ \\
		\cline{2-4} & 3.68 & 0.00 & $\times$ \\
		\cline{2-4} & 5.43 & 0.00 & $\times$ \\
		\hline
		\multirow{3}*{1.49} & 4.76 & $0.00$ & $\times$ \\
		\cline{2-4} & 7.27 & $0.00$ & $\times$ \\
		\cline{2-4} & 1.52 & $0.00$ & $\times$ \\
		\hline
	\end{tabular}
\end{table}
